Inefficient Concessions and Mediation

### 2023-03-20 draft_narrow.tex
misc nits

In efforts to explain the existence of gift giving,  Camerer (1988) and Van de Ven (2000) give several anthropological explanations for inefficiency. Amongst these explanations are altruism, social mores, and egoism. But the most relevant explanation for the field of international relations is that gift giving is strategic. Camerer uses mechanism design to model gifts as costly signals; inefficiency is useful in pre-play communication in order to form relationships between like types. This strategic giving of gifts in societies looks very much like the strategic giving of concessions to remedy conflict situations. When modeled game-theoretically, these environments look even more similar.

@@

Prendergast and Stole (2001) provide a second explanation for inefficient gifts. The authors explain inefficient gift giving by modeling the utility of matching. That is, if players are of a like mindset they get an extra benefit so they are willing to take the risk of giving the wrong gift.

We propose a different conceptual framework to explain this real-life phenomenon that is rooted in a realistic international relations puzzle. The basic formulation of our mechanism-design approach is an expansion of Camerer's (1988) analysis. The model can achieve the basic `inefficient gifts' results of Camerer (1988) and Prendergast and Stole (2001) but does so using a very different explanation for inefficiency. The basic model is a costly signaling model, at its root similar to Spence (1973) and subsequent papers.\footnote{See Connelly et al. (2011) for a review of the costly signaling literature.} In essence, the model shows that if concessions can be later used against the party that gives them, then it can be in a negotiating party's best interest to give inefficient concessions. While our model is tailored to an international relations setting, the explanation for inefficient giving of concessions may be more broadly applicable.

@@

\subsection{Pooling equilibrium}
In the \emph{pooling equilibrium}, neither negotiating party gives concessions in period 0, then both parties play Fight in every period beginning with period 1. Countries do nothing to bridge the credibility gap, and thus do not cooperate. This \emph{pooling equilibrium} is likely to happen when countries are in long-set patterns of distrust. Active fighting is not necessary; remaining outside of a state of peace is all that is required.  See the Appendix for a formal characterization of this type of equilibrium.

Cyprus' civil war is a good example of the pooling equilibrium where neither side is taking actions to move things forward. While both sides posture that they want peace, and while there has been some lessening of tensions, very few concrete actions have been taken that move the sides toward settlement. This conflict is a stalemate.

Sri Lanka is another example. While the government has finally won the war of secession, it still refuses to address grievances of the minorities and leaves the conflict unresolved. Other prominent examples of this pooling equilibrium include Israel and its neighbors, North and South Korea, Greece and Turkey, and a multitude of other long-lasting conflicts. The mistrust is perhaps justified; indeed if one party did take the first step and give a concession first, they may be taken advantage of.

>>>>

In efforts to explain the existence of gift giving,  Camerer (1988) and Van de Ven (2000) give several anthropological explanations for inefficiency.\footnote{Prendergast and Stole (2001) also provide an explanation for inefficient gift giving focused on the utility of matching.} Amongst these explanations are altruism, social mores, and egoism. But the most relevant explanation for the field of international relations is that gift giving is strategic. Camerer uses mechanism design to model gifts as costly signals; inefficiency is useful in pre-play communication in order to form relationships between like types. This strategic giving of gifts in societies looks very much like the strategic giving of concessions to remedy conflict situations. When modeled game-theoretically, these environments look even more similar.

@@

%Prendergast and Stole (2001) provide a second explanation for inefficient gifts. The authors explain inefficient gift giving by modeling the utility of matching. That is, if players are of a like mindset they get an extra benefit so they are willing to take the risk of giving the wrong gift.

We propose a different conceptual framework to explain this real-life phenomenon that is rooted in a realistic international relations puzzle. The basic formulation of our mechanism-design approach is an expansion of Camerer's (1988) analysis. The model can achieve the basic `inefficient gifts' results of Camerer (1988) %and Prendergast and Stole (2001) 
but does so using a very different explanation for inefficiency. The basic model is a costly signaling model, at its root similar to Spence (1973) and subsequent papers.\footnote{See Connelly et al. (2011) for a review of the costly signaling literature.} In essence, the model shows that if concessions can be later used against the party that gives them, then it can be in a negotiating party's best interest to give inefficient concessions. While our model is tailored to an international relations setting, the explanation for inefficient giving of concessions may be more broadly applicable.

@@

\subsection{Pooling equilibrium}
In the \emph{pooling equilibrium}, neither negotiating party gives concessions in period 0, then both parties play Fight in every period beginning with period 1. Countries do nothing to bridge the credibility gap, and thus do not cooperate. This \emph{pooling equilibrium} is likely to happen when countries are in long-set patterns of distrust. Active fighting is not necessary; remaining outside of a state of peace/cooperation is all that is required.  See the Appendix for a formal characterization of this type of equilibrium.

%Cyprus' civil war is a good example of the pooling equilibrium where neither side is taking actions to move things forward. While both sides posture that they want peace, and while there has been some lessening of tensions, very few concrete actions have been taken that move the sides toward settlement. This conflict is a stalemate.

Sri Lanka offers one example. While the government has finally won the war of secession, it still refuses to address grievances of the minorities and leaves the conflict unresolved. Other prominent examples of this pooling equilibrium include Israel and its neighbors, Cyprus's civil war, North and South Korea, Greece and Turkey, and a multitude of other long-lasting conflicts. The mistrust is perhaps justified; indeed if one party did take the first step and give a concession first, they may be taken advantage of.

Outside the war context, [EXAMPLE NEEDED].
